---
title: "Uber Pickups in New York City"
author: "Williams"
date: "6/5/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepartion

```{r, message=FALSE}
##Importing the packages
library(animation)
library(ggpmisc)
library(reshape2)#Melt function
library(mice)##checking NA data
library(ggmap)###google map
library(stringi)#interval function
library(lubridate)###Date.Time Format
library(dplyr)#group function
library(base)
library(anytime) #month function
library(ggplot2)####
library(tidyverse)
library(knitr)
library(ggthemes)
library(readr)
library(viridis)
library(hexbin)
library(scales)
library(plotly)
library(kableExtra)
library(gganimate)
library(corrplot)
library(GGally)
library(tmap)
```

```{r, message=FALSE}
##Reading the files
apr14 <- read_csv("project/archive (5)/uber-raw-data-apr14.csv")
may14 <- read_csv("project/archive (5)/uber-raw-data-may14.csv")
jun14 <- read_csv("project/archive (5)/uber-raw-data-jun14.csv")
jul14 <- read_csv("project/archive (5)/uber-raw-data-jul14.csv")
aug14 <- read_csv("project/archive (5)/uber-raw-data-aug14.csv")
sep14 <- read_csv("project/archive (5)/uber-raw-data-sep14.csv")
```

```{r, message=FALSE}
##Merging all the data in 2014
rbind(apr14, may14, jun14, jul14, aug14, sep14) -> uber_raw_date
uber_total <- uber_raw_date
```

# Data cleaning and formatting

```{r, message=FALSE}
#Changing Date.Time format
uber_total$`Date_Time` <- as.POSIXct(uber_total$`Date/Time`, format = "%m/%d/%Y %H:%M:%S")
```

```{r, message=FALSE}
#Adding 'Month','Day','Hour','Weekday' column, changing the date to the right format
uber_total %>%
     mutate(Date_Time = as.POSIXct(`Date/Time`, format = "%m/%d/%Y %H:%M:%S"),
            Month = month(Date_Time),
            Day = day(Date_Time),
            Hour = hour(Date_Time),
            Weekday = wday(Date_Time),
            Date = date(Date_Time)) -> uber_total
```

# Data analysis and visualisation

```{r, message=FALSE}
#Trips per month
uber_total %>%
  group_by(Month) %>%
  tally() %>%
  ggplot(aes(x= Month , y= n)) + 
  scale_x_continuous(breaks = seq(4,9,1)) +
  geom_col(fill="blue")+
  labs(x= 'Month' ,
       y= 'Trips' ,
       title = 'Trips each month in 2014' ,
       subtitle = 'There was a big rise in September') +
  scale_y_continuous(labels = comma)
```

```{r}
#Trips per hour
uber_total %>%
  group_by(Hour) %>%
  tally()%>%
  ggplot(aes(x= Hour , y= n)) + 
  geom_col(fill="red")+
  scale_x_continuous(breaks = seq(0,23,1)) +
  labs(x= 'Hour' ,
       y= 'Trips' ,
       title = 'Trips each hour in 2014' ,
       subtitle = 'Peak hours in 7-8 AM and 4-6 PM') +
  scale_y_continuous(labels = comma)
```

```{r, message=FALSE}
#Adding day_of_the_week column
uber_total %>%
  mutate(
    day_of_the_week = case_when(
    Weekday == 1 ~ 'Sunday' ,
    Weekday == 2 ~ 'Monday' ,
    Weekday == 3 ~ 'Tuesday' ,
    Weekday == 4 ~ 'Wednesday' ,
    Weekday == 5 ~ 'Thursday' , 
    Weekday == 6 ~ 'Friday' ,
    Weekday == 7 ~ 'Saturday'
  )
  ) -> uber_total
```

```{r, message=FALSE}
#Adjusting week of the day to the right order for plots below
uber_total$day_of_the_week <- factor(uber_total$day_of_the_week, levels=c("Sunday","Saturday","Friday","Thursday", "Wednesday","Tuesday","Monday"))
```


```{r, message=FALSE}
#Hourly pickups by day of Week in 2014
uber_total %>%
  group_by(Hour,day_of_the_week) %>%
  tally()%>%
  ggplot(mapping = aes(x=Hour, y=day_of_the_week, fill=n))+
  geom_tile(colour='white') +
  scale_x_continuous(breaks = seq(0,23,1)) +
  theme_minimal()+
  scale_fill_viridis(option = 'D') +
  labs(x= 'Hour' ,
       y= 'Weekday' ,
       subtitle = 'Hourly pickups by day of Week in 2014' ,
       fill = 'Number of pickups')
```

```{r, message=FALSE}
weekdaylist<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
uber_total$Workday=ifelse(uber_total$day_of_the_week %in% weekdaylist ,"Weekday","Weekend")
```

```{r, message=FALSE}
uber_total %>%
  group_by(Date,Workday)%>%
  summarise(Trips = n())%>% 
  ggplot(mapping = aes(x=Date,y=Trips,color=Workday))+
  scale_color_manual(values =c("royalblue4","red2"))+
  geom_point()+geom_smooth(method=lm, fill='yellowgreen', 
                           color='yellow1',se=FALSE,size=1, formula = y~x)+
  ggtitle("Pickups:Weekdays vs. Weekends")+ylab("Total Pickups per Day")
```


```{r, message=FALSE}
#Pickup numbers change monthly per base
uber_total %>%
  group_by(Month, Base) %>%
  tally()%>%
  ggplot(mapping = aes(x=Month, y=n, colour=Base))+
  geom_line()+
  theme_minimal()+
  scale_y_continuous(labels = comma)+
  labs(x= 'Month',
       y = 'Pickup',
       title = 'Pickup numbers change monthly per base',
       subtitle = 'B02764 and B02617 see big rise since August')
```

```{r, message=FALSE}
#Hex Map of Geographical Location of Uber Rides
ggplot(data = uber_total , mapping= aes(x = Lat, y = Lon)) + 
    geom_hex(bins = 80) + 
    ggtitle('Hex Map of Geographical Location of Uber Rides') + 
    xlab('Latitude (Degrees North)')  +
    ylab('Longitude (Degrees East)') +
  labs(subtitle = 'Most rides located around 40.7° N, 74° W')
```

```{r, message=FALSE}
#NY Uber Rides Map During 2014 Apr-Sep
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004
ggplot(uber_total, aes(x=Lon, y=Lat)) +
geom_point(size = 0.1 , colour= 'blue') +
scale_x_continuous(limits = c(min_long, max_long)) +
scale_y_continuous(limits = c(min_lat, max_lat)) +
theme_map()  +
ggtitle("NY Uber Rides Map During 2014 Apr-Sep")+
labs(subtitle = 'Most rides located at Manhattan, Brooklyn, Queens')
```

```{r, message=FALSE}
#Choosing the Top 30 locations in 2014 by total volumes
library(leaflet)
uber_2014_top <- uber_total %>%
    count(Lat, Lon, sort = TRUE) %>%
    head(30)
```

```{r, message=FALSE}
#The Top 30 locations in 2014 by total volumes map
leaflet(uber_2014_top) %>%
  addTiles() %>%
  setView(-74.15, 40.91, zoom = 8) %>%
  addMarkers(~Lon, ~Lat)
```

```{r, message=FALSE}
#Choosing the Top 30 locations in 2014 by total volumes in weekday morning peak hours
uber_morning_peak_weekday <- uber_total %>%
    filter(Workday == 'Weekday') %>%
    filter(Hour %in% c(7,8,9)) %>%
    count(Lat, Lon, sort = TRUE) %>%
    head(30)
    
```

```{r, message=FALSE}
#The Top 30 locations in 2014 by total volumes map in weekday morning peak hours
leaflet(uber_morning_peak_weekday) %>%
    addTiles() %>%
    setView(-74.15, 40.91, zoom = 8) %>%
    addMarkers(~Lon, ~Lat)
```

```{r, message=FALSE}
#Choosing the Top 30 locations in 2014 by total volumes in weekday evening peak hours
uber_evening_peak_weekday <- uber_total %>%
    filter(Workday == 'Weekday') %>%
    filter(Hour %in% c(17,18,19,20)) %>%
    count(Lat, Lon, sort = TRUE) %>%
    head(30)
```

```{r, message=FALSE}
#The Top 30 locations in 2014 by total volumes map in weekday evening peak hours
leaflet(uber_evening_peak_weekday) %>%
    addTiles() %>%
    setView(-74.15, 40.91, zoom = 8) %>%
    addMarkers(~Lon, ~Lat)
```

```{r, message=FALSE}
#Choosing the Top 30 locations in 2014 by total volumes in weekend party time
uber_party_1 <- uber_total %>%
    filter(Workday == 'Weekend') %>%
    filter(Hour %in% c(0,1,2,3)) %>%
    count(Lat, Lon, sort = TRUE) %>%
    head(30)
```

```{r, message=FALSE}
#The Top 30 locations in 2014 by total volumes map in weekend party time
leaflet(uber_party_1) %>%
    addTiles() %>%
    setView(-74.15, 40.91, zoom = 8) %>%
    addMarkers(~Lon, ~Lat)
```

```{r, message=FALSE}
#Another version of party map
#To get the map of NY City
newyork <- get_map(location = c(-74.5477,40.4553,-73.1525,41.0618),source = "osm")
```


```{r, message=FALSE}
#Filtering the data of party time
uber_party_2 <- uber_total %>%
    filter(Workday == 'Weekend') %>%
    filter(Hour %in% c(0,1,2,3))
```

```{r, message=FALSE, warning=FALSE}
#Getting the map
ggmap(newyork)+
  geom_point(data = uber_party_2, mapping = aes(x= Lon , y= Lat) ,alpha = 0.03)
```

```{r, message=FALSE, warning=FALSE}
#The Gif of NY Uber Rides Map Changes During Apr 2014
p <- ggplot(data=subset(uber_total,Month==4), aes(x=Lon, y=Lat)) +
geom_point(size = 0.1 , colour= 'blue') +
scale_x_continuous(limits = c(min_long, max_long)) +
scale_y_continuous(limits = c(min_lat, max_lat)) +
theme_map()  +
labs(subtitle = 'Date:{frame_time}')+
ggtitle("NY Uber Rides Map Changes During Apr 2014")+
transition_time(Date)
```


```{r, message=FALSE, warning=FALSE}
p
```

# I decided to add data in 2015 to have a deep understanding in NY Uber rider situation

```{r, message=FALSE}
#Changing the format of the original table to merge data in 2 years
uber_total %>% select(`Date_Time`,'Base') -> uber14
names(uber14) <- c("Date","Base")
```


```{r, message=FALSE}
#Loading the data in 2015 and changing the format
janjune_15 <- read_csv("project/archive (5)/uber-raw-data-janjune-15.csv")
janjune_15 %>% select(`Pickup_date`,'Affiliated_base_num') -> uber15
names(uber15) <- c("Date","Base")
```

```{r, message=FALSE}
##Merging the data in 2014 and 2015
uber_all <- bind_rows(uber14, uber15)
```

```{r, message=FALSE}
#Changing date format
uber_all$Date <- as.POSIXct(uber_all$Date, format = "%Y-%m-%d %H:%M:%S")
```


```{r, message=FALSE}
#Adding 'year', 'month','day','weekday','hour','minute' column, changing the date to the right format
uber_all %>% mutate(year = year(Date),
                            month = month(Date),
                            monthname = month(Date, label = TRUE, abbr = FALSE),
                            day = day(Date),
                            weekday = wday(Date, label = TRUE, abbr = FALSE),
                            hour = hour(Date),
                            minute = minute(Date)) -> uber_all
```

```{r, message=FALSE}
#Picking data only in April, May ,June 2015 to compare situations in these 3 months in 2014 and 2015
uber_all %>% 
  filter(month == 4 | month == 5 | month == 6) -> uber_all
```


```{r, message=FALSE}
#Adjusting week of the day to the right order for plots below
uber_all$weekday <- factor(uber_all$weekday, levels=c("Sunday","Saturday","Friday","Thursday", "Wednesday","Tuesday","Monday"))
```

```{r, message=FALSE}
#Hourly pickups by day of Week in 2014
uber_all %>%
  filter(year == 2015)%>%
  group_by(weekday, hour)%>%
  summarise(Trips = n())%>%
  ggplot(mapping = aes(x=hour, y=weekday, fill=Trips))+
  geom_tile(colour='white') +
  scale_x_continuous(breaks = seq(0,23,1)) +
  theme_minimal()+
  scale_fill_viridis(option = 'D') +
  labs(x= 'Hour' ,
       y= 'Weekday' ,
       subtitle = 'Hourly pickups by day of Week in 2015' ,
       fill = 'Number of pickups')
```

```{r, message=FALSE}
#Trips per hour
uber_all %>%
  group_by(hour) %>%
  summarise(Trips = n())%>%
  ggplot(aes(x= hour , y= Trips)) +
  scale_x_continuous(breaks = seq(0,23,1)) +
  geom_col(fill="red")+
  labs(x= 'Hour' ,
       y= 'Trips' ,
       title = 'Trips each hour in 2015 Apr-Jun' ,
       subtitle = 'Peak hours in 8-10 AM and 6-11 PM') +
  scale_y_continuous(labels = comma)
```

```{r, message=FALSE}
#The Volume Difference By Bases Between 2014 And 2015
uber_all %>%
  group_by(year,Base) %>%
  filter(Base %in% c('B02512','B02598','B02617','B02682','B02764'))%>%
  summarise(Trips = n())%>%
  ggplot(mapping = aes(x=reorder(Base,Trips),y=Trips,colour=as.factor(year)))+
  geom_point()+
  coord_flip()+
  scale_color_discrete(name = "Year:")+
  theme_classic()+
  labs(y= 'Base' ,
       title = 'The Volume Difference By Bases Between 2014 And 2015' ,
       subtitle = 'A big rise in B02682 and B02764 after one year')
```

# Conclusion
#### 1.Most trips happened at 7-8 AM at weekday morning, 6-10 PM at weekday evening , 0-2 AM at weekend days.
#### 2.Most trips located at two airports(John F. Kennedy International Airport and LaGuardia Airport)
#### 3.Most trips located at Manhattan, Brooklyn and Queens


